Roster Review (Generic)

### 2019-07-09 rosterReview_generic.Rmd
optimal lineup complete, experimenting with various ros projected team rankings and views

```{r}
force_fetch <- FALSE
force_join <- FALSE
batter_file <- "temp/FanGraphs Leaderboard.csv"
pitcher_file <- "temp/FanGraphs Leaderboard (1).csv"

@@

```{r}
rosters_by_position <- row_per_position(rosters)

@@

# TODO get position factor ordering back, summarize pitching to single point/range, add bench single point/range, consider DL?
rosters_by_position %>% 
  filter(team != "Free Agent",
         type == "pitcher") %>% 
  mutate(position = as.character(position)) %>% 
  droplevels() -> team_pitchers_all

@@

team_pitchers_RP <- team_pitchers_all %>% 
  filter(position == 'RP') %>% 
  anti_join(team_pitchers_SP, "player_id") %>% 
  group_by(team) %>% 
  arrange(zar %>% desc) %>% 
  mutate(rnk = row_number()) %>% 
  filter(rnk <= 6) %>% 
  select(position, player_id, name_full, team, manager, zar, pos_zar)

team_pitchers_SP %>% 
  bind_rows(team_pitchers_RP) %>% 
  group_by(team, position) %>% 
  summarize(zar = mean(zar),
            pos_zar = mean(pos_zar)) -> team_pitchers_starters

@@

team_batters_starters %>% 
  bind_rows(team_batters_bench) %>% 
  bind_rows(team_pitchers_starters)

#FIXME team_batters_starters needs MANAGER column

@@

```{r}
#TODO abstract some of these views
rosters %>% 
  mutate(zar = round(zar)) %>% 
  filter((team == 'Free Agent' & zar >=0) | (grepl("Plouffe",team) & zar < 2)) %>% 
  ggplot(aes(x=team,y=zar,color=type)) +
  geom_jitter(aes(text=glue("{name_full} - {display_position}", color=type))) +
  coord_flip() -> pa

@@

```{r}
rosters_by_position %>% 
  select(team, manager, name_full, position, zar, pos_zar) %>% 
  # filter(position %in% c('SP','RP')) %>% 
  filter(grepl("Plouffe",team)) %>% 
  arrange(pos_zar %>% desc) %>%
  mutate(pos_zar = pos_zar %>% round(1),
         zar = zar %>% round(1)) %>% 
  View("ZAR by Position")

rosters_by_position %>% 
  filter(team != "Free Agent") %>% 
  ggplot(aes(position, pos_zar, color = position, alpha = zar)) +
  geom_point(aes(text=name_full)) +
  facet_wrap(~ team) -> team_pos_adj

team_pitchers %>% 
  bind_rows(team_batters_starters) %>% 
  ggplot(aes(position, pos_zar, color = position, alpha = zar)) +
  geom_point() +
  facet_wrap(~ team) -> team_pos_optimal

team_pos_avg_zar %>% 
  ungroup() %>% 
  group_by(team) %>% 
  summarize(sum_pos_zar = sum(avg_pos_zar),
            mean_pos_zar = mean(avg_pos_zar),
            var_pos_zar = sd(avg_pos_zar)) %>% 
  arrange(mean_pos_zar %>% desc)
```

```{r}
# TODO find way to wrap these three views into one
rosters %>% 
  filter(grepl("Plouffe|Mata",team)) %>% 
  select(team, manager, name_full, display_position, zar, positions) %>% 
  arrange(zar %>% desc) %>% 
  mutate(zar = zar %>% round(1)) %>% 
  View("Trade Partner")
  # mutate(zar = janitor::round_half_up(zar)) %>% 
  # write_csv("data/tradeReview_20190423.csv")

@@

```{r}
rosters %>% 
  filter(!grepl("Free Agent",team)
  #        ,
  # grepl("CF",positions)
       ) %>%
  ggplot(aes(x=team,y=zar,color=type)) +
  geom_boxplot() +
  # geom_jitter(aes(text=name_full), width = .2) +
  coord_flip() -> tb

ggplotly(tb)
```

```{r}
#TODO measuring projected value of draft picks, move to separate research
#TODO get beginning of year zar/war values
rosters %>% 
  filter(!is.na(draft_pick)) %>%
  group_by(draft_round) %>% 
  summarize(zar_median = median(zar) %>% janitor::round_half_up(),
            zar_mean = mean(zar) %>% janitor::round_half_up()) %>%
  ggplot(aes(x=draft_round,y=zar_mean)) +
  geom_col() +
  geom_line(aes(y=zar_median)) -> draft_plot

>>>>

```{r get_data}
force_fetch <- FALSE
force_join <- FALSE
batter_file <- "temp/FanGraphs Leaderboard.csv"
pitcher_file <- "temp/FanGraphs Leaderboard (1).csv"

@@

```{r df_prep}
rosters_by_position <- row_per_position(rosters)

@@

# TODO get position factor ordering back, consider DL?
rosters_by_position %>% 
  filter(team != "Free Agent",
         type == "pitcher") %>% 
  mutate(position = as.character(position)) %>% 
  droplevels() -> team_pitchers_all

@@

team_pitchers_RP <- team_pitchers_all %>% 
  filter(position == 'RP') %>% 
  anti_join(team_pitchers_SP, "player_id") %>% 
  group_by(team) %>% 
  arrange(zar %>% desc) %>% 
  mutate(rnk = row_number()) %>% 
  filter(rnk <= 6)

team_pitchers_SP %>% 
  bind_rows(team_pitchers_RP) %>% 
  select(position, player_id, name_full, team, manager, zar, pos_zar) -> team_pitchers_starters

@@

team_lineups <- team_batters_starters %>% 
  bind_rows(team_batters_bench) %>% 
  bind_rows(team_pitchers_starters) %>% 
  mutate(position = fct_relevel(position, "C","1B","2B","3B","SS","LF","CF","RF", "Util","SP","RP","BN"))

standings_projected <- team_lineups %>% 
  filter(position != 'BN') %>% 
  group_by(team) %>% 
  summarize(sum_pos_zar = sum(pos_zar),
            mean_pos_zar = mean(pos_zar),
            median_pos_zar = median(pos_zar),
            var_pos_zar = sd(pos_zar)) %>% 
  arrange(sum_pos_zar %>% desc)

@@

```{r ros_team_rankings}
rosters_by_position %>% 
  filter(team != "Free Agent") %>% 
  ggplot(aes(position, pos_zar, color = position, alpha = zar)) +
  geom_point(aes(text=name_full)) +
  facet_wrap(~ team) -> team_pos_adj

team_lineups %>% 
  ggplot(aes(position, pos_zar, color = position, alpha = zar)) +
  geom_point(aes(text=name_full)) +
  facet_wrap(~ team) -> team_pos_optimal

standings_projected %>% 
  mutate(sum_pos_zar = sum_pos_zar - median(sum_pos_zar)) %>% 
  mutate(team = fct_reorder(team, sum_pos_zar)) %>% 
  ggplot(aes(team, sum_pos_zar)) +
  geom_col() +
  coord_flip()

#TODO try other graph views, including error bars representing variance
```

```{r free_agents}
#TODO abstract some of these views
rosters %>% 
  mutate(zar = round(zar)) %>% 
  filter((team == 'Free Agent' & zar >=0) | (grepl("Plouffe",team) & zar < 2)) %>% 
  ggplot(aes(x=team,y=zar,color=type)) +
  geom_jitter(aes(text=glue("{name_full} - {display_position}", color=type))) +
  coord_flip() -> pa

@@

```{r trade_eval}
# TODO find way to wrap these three views into one
rosters %>% 
  filter(grepl("Plouffe|Mata",team)) %>% 
  select(team, manager, name_full, display_position, zar, positions) %>% 
  arrange(zar %>% desc) %>% 
  mutate(zar = zar %>% round(1)) %>% 
  View("Trade Partner")
  # mutate(zar = janitor::round_half_up(zar)) %>% 
  # write_csv("data/tradeReview_20190423.csv")

@@

```{r draft_analysis}
#TODO measuring projected value of draft picks, move to separate research
#TODO get beginning of year zar/war values
rosters %>% 
  filter(!is.na(draft_pick)) %>%
  group_by(draft_round) %>% 
  summarize(zar_median = median(zar) %>% janitor::round_half_up(),
            zar_mean = mean(zar) %>% janitor::round_half_up()) %>%
  ggplot(aes(x=draft_round,y=zar_mean)) +
  geom_col() +
  geom_line(aes(y=zar_median)) -> draft_plot
